\documentclass[12pt]{report}
\usepackage{consumer_resource_final}
\begin{document}
\subsection{Why do we solve it this way}\label{sec: explanation solve for d and m}

\subsection{Algorithmic procedure}\label{sec : feasibility methods algorithmic procedure}
We hereby detail the $\mathcal{A}$ procedure which from a set of metaparameters $m$ and a consumption-syntrophy network $(G,A)$ gives rise to a set of parameters $p$. It goes like this:
\begin{enumerate}
  \item We first draw randomly each $R^*_\nu$
  and $S^*_i$ :
  \begin{equation}
     R^*_\nu = \mathcal{R} \ \forall \nu=1, \dots, N_R\text{ and }  S^*_i = \mathcal{S} \ \forall i=1, \dots, N_S,
  \end{equation}
  where $\mathcal{R}$ and $\mathcal{S}$ are random variables coming from a distribution of mean equal to the corresponding metaparameter and relative standard deviation\footnote{By relative standard deviation, we mean the standard deviation measured in units of the average value.} $\epsilon$. In our simulations, we chose uniform distributions :
  \begin{equation}
  \mathcal{R} \sim \text{Unif}(R_0, R_0 \epsilon) \text{ and } \mathcal{S} \sim \text{Unif}(S_0, S_0\epsilon).
  \end{equation}
  \item The efficiency matrix $\sigma_{i\nu}$ is then drawn similarly, from a distribution with average $\sigma_0$. In order to simplify the problem\footnote{Indeed, we observerd that in general introducing a non uniform $\sigma_0$ adds an additional not needed layer of complexity.}, we take a zero-variance like \citeauthor{butler_stability_2018} in \cite{butler_stability_2018}, \ie all species consume resources at the same global efficiency :
  \begin{equation}
    \sigma_{i\nu} = \sigma_0.
  \end{equation}
  \item We build the consumption matrix $\gamma_{i\nu}$. Its adjacency matrix $G$ is loaded through a user-provided file.
  %$G$ is a binary matrix given by the user (in the \code{configuration.in} file) and is defined as the adjacency matrix of the consumption network (\ie it tells which species eats which resource). We then build $\gamma$ with the same network structure as $F$ (\ie both matrices have the same zero elements).
  While $G$ gives the structure of $\gamma$, \ie if a given $\gamma_{i\nu}$ is zero or not, the actual values of $\gamma_{i\nu}$ need then to be determined. They are drawn from a uniform distribution of mean $\gamma_0$ and relative standard deviation $\epsilon$:
  \begin{equation}
    \gamma_{i\nu} = \text{Unif}(\gamma_0, \gamma_0\epsilon) G_{i\nu}.
  \end{equation}
  \item We draw the resources external feeding rates, similarly to the other parameters :
  \begin{equation}
  l_\mu = \text{Unif}(l_0, l_0\epsilon) \ \forall \mu=1, \dots, N_R.
  \end{equation}
  \item The last free parameters are the non-zero elements of the syntrophy matrix $\alpha_{\nu i}$, since the $d_i$ and $l_\mu$ are determined through the equations of evolution at equilibrium. %This is the tricky part of the algorithm because $\alpha$ has to follow three constraints, namely energy conservation/dissipation Eq.\eqref{eq : dissipation constraint} and positiveness of $d_i$ and $l_\mu$ [\textbf{insert reference to equation}]. The general strategy is to choose the metaparameters in a way that these constraints should \textit{almost always} be satisfied, \ie we pick metaparameters that follow the feasibility constraint Eq.\eqref{eq : overall feasibility constraint metaparameters}.
  The adjacency matrix $A$ of $\alpha$ needs then to be specified. At the moment, it can be chosen in four different ways : fully connected, in a way that no resource eaten by a given species can be released by that same species (\ie no intraspecific syntrophy : $G_{i\mu}>0 \iff A_{\mu i}=0$), random structure with the same number of links as the consumption matrix and finally by a user provided matrix. After the adjacency matrix is loaded, we build $\alpha$ from a uniform distribution of mean $\alpha_0$ and relative standard deviation $\epsilon$:
  \begin{equation}
    \alpha_{\nu i} = \text{Unif}(\alpha_0, \alpha_0\epsilon) A_{\nu i} .
  \end{equation}
  %\item We build $\tau_{\nu i}$. It usually is equal to $\alpha_{\nu i}$ or 0.
  \item With all of these parameters drawn, we can solve Eqs.\eqref{eq : feasibility positive d} for the species death rate $d_i$ and Eqs.\eqref{eq : feasibility positive m} for $m_\nu$. All the parameters of the model are now fully determined.
  \item We check if the constraints Eqs.\eqref{eq : feasibility positive d and m} and \eqref{eq : feasability biomass conservation} on the parameters are fulfilled. If they are not, we go back to step 1. Otherwise, we can exit the algorithm, a feasible system has been built. We see here the advantage of having input metaparameters that will most likely give rise to a feasible parameter set. If we just give random metaparameters, we run the risk of getting stuck in an unpredictably long loop between steps 1 and 7. If however we are in a region where we know the metaparameters feasibility is high, a feasible system is found much faster.
\end{enumerate}

\subsection{When is zero part of the spectrum of $J^*$?}\label{sec : zero part of spectrum}
We are interested in knowing when $\lambda=0$ is part of the spectrum of $J^*$. By definition, $\lambda=0$ is an eigenvalue if and only if it solves the master equation \eqref{eq: master equation eigenvalues jacobian}:
\begin{equation}
\det
\begin{pmatrix}
  -\Delta   & \Gamma \\
  \Beta & 0
\end{pmatrix} = 0 \label{eq: determinant marginally stable equilibria}
\end{equation}
Using the fact that $\Delta$ is invertible, we can make use of the equality\footnote{This uses a formula which is trivially analogous to one found in \cite{powell_calculating_2011}.}:
\begin{equation}
\det\begin{pmatrix}
  -\Delta   & \Gamma \\
  \Beta & 0
\end{pmatrix} = \det(-\Delta)\det(\Beta\Delta^{-1}\Gamma).
\end{equation}
Eq.\eqref{eq: determinant marginally stable equilibria} then becomes:
\begin{equation}
\det(\Beta\Delta^{-1}\Gamma)=0
\end{equation}
which means that $\Beta \Delta^{-1}\Gamma$ is not full rank. Finally,
\begin{equation}
\boxed{
0 \in \sigma(J^*) \iff \Beta \Delta^{-1}\Gamma \text{ is not full rank.}\label{eq : condition zero in spectrum of J}
}
\end{equation}
But when is $\Beta \Delta^{-1}\Gamma$ not full rank? Sylvester rank inequality \cite{thome_inequalities_2016} states that:
\begin{equation}
\text{rank}(\Beta\Delta^{-1}\Gamma) \geq \text{rank}(\Beta)+\text{rank}(\Delta^{-1}\Gamma)
-N_R.
\end{equation}
Similarly,
\begin{equation}
\text{rank}(\Delta^{-1}\Gamma) \geq \text{rank}(\Delta^{-1})+\text{rank}(\Gamma)-N_R=\text{rank}(\Gamma),
\end{equation}
where we used the fact that $\Delta^{-1}$ is invertible so $\text{rank}(\Delta^{-1})=N_R$.
One of the standard rank properties is:
\begin{equation}
\text{rank}(\Delta^{-1}\Gamma) \leq \min\left\{\text{rank}(\Delta^{-1}), \text{rank}(\Gamma)\right\} \implies \text{rank}(\Delta^{-1}\Gamma) \leq \text{rank}(\Gamma).
\end{equation}
In the end this yields $\text{rank}(\Delta^{-1}\Gamma)=\text{rank}(\Gamma)$ and :
\begin{equation}
\text{rank}(\Beta\Delta^{-1}\Gamma) +N_R \geq \text{rank}(\Beta)+\text{rank}(\Gamma). \label{eq : lower bound Beta Delta Gamma}
\end{equation}
We can use this inequality to enunciate a lemma about the presence of zero in the spectrum of $J^*$.
\lemma{If $N_R \leq N_S$ and $\Beta$ and $\Gamma$ are full rank, then $0 \notin \sigma(J^*)$.}
\begin{proof}
{We assume that $0$ is in the spectrum of $J^*$ and prove it leads to a contradiction. Because $0 \in \sigma(J^*)$, Eq.\eqref{eq : condition zero in spectrum of J} implies $\Beta\Delta^{-1}\Gamma$ is not full rank. Since the largest possible value for the rank of matrix is the minimum between its number of rows and columns, we have :
\begin{equation}
\text{rank}(\Beta\Delta^{-1}\Gamma)< \min(N_R, N_S) = N_R.
\end{equation}
This can be used as an upper bound for Eq.\eqref{eq : lower bound Beta Delta Gamma} :
\begin{equation}
2 N_R > \text{rank}(\Beta)+\text{rank}(\Gamma).
\end{equation}
However, we also know that $\Beta$ and $\Gamma$ are full rank, \ie
\begin{equation}
\text{rank}(\Beta)=\text{rank}(\Gamma)=N_R.
\end{equation}
Hence the previous inequality amounts to $N_R > N_R$, which is a contradiction. We conclude that the hypothesis $0 \in \sigma(J^*)$ is wrong.
}
\end{proof}

\subsection{Weak LRI regime}\label{sec : weak LRI regime}
\begin{theorem}
Let $p$ be a parameter set with a jacobian at equilibrium $J^*$. If $0$ is not an eigenvalue of $J^*$ and the equations
\begin{equation}
\left(\Gamma \Beta\right)_{\mu\mu} < - \sum_{\nu \neq \mu } \abs{\left(\Gamma \Beta\right)_{\mu \nu}} \ \forall \mu,
\end{equation}
are verified, then the real eigenvalues of $J^*$ are negative.
\end{theorem}
\begin{proof}
We assume
\begin{equation}
\left(\Gamma \Beta\right)_{\mu\mu} < - \sum_{\nu \neq \mu } \abs{\left(\Gamma \Beta\right)_{\mu \nu}}  \ \forall \mu.
\end{equation}
Let $\lambda \in \mathbb{R}$, then the following will also trivially hold:
\begin{equation}
  \left(\Gamma\Beta \right)_{\mu\mu} -\lambda^2 < - \sum_{\nu \neq \mu } \abs{\left(\Gamma \Beta\right)_{\mu \nu}} \ \forall \mu. \label{eq: low species bound 1}
\end{equation}
Dividing Eq.\eqref{eq: low species bound 1} by $\Delta_\mu$, we get:
\begin{equation}
\frac{1}{\Delta_\mu}\left[\left(\sum_i \Gamma_{\mu i} \Beta_{i \mu} \right)-{\lambda^2}\right] < - \sum_{\nu \neq \mu } \abs{\frac{\sum_i \Gamma_{\mu i}\Beta_{i\nu}}{\Delta_\mu}} \ \forall \mu.
\end{equation}
Looking at Eq.\eqref{eq: definition S component wise}, we see that this is equivalent to:
\begin{equation}
S_{\mu \mu} + \sum_{\nu \neq \mu} \abs{S_{\mu \nu}} < 0\ \forall \mu.
\end{equation}
Using Lemma \ref{lemma: lemma Gerschgorin circle}, we know that all the real eigenvalues of $S(\lambda)$ will have a negative real part.
We can conclude with the statement of the theorem.
\end{proof}


\subsection{Effective system}
Models which involve the dynamics of species only are in general better known than consumers-resources models [\textbf{insert reference}]. In particular, a huge body of literature exists on the study of Lotka-Volterra systems [\textbf{insert reference}]. We may profit from this knowledge by transforming the effect of the resources dynamics into an effective consumers-only system.

This can be done by assuming that the resources reach an equilibrium way faster than the consumers.
Mathematically, that is equivalent to
\begin{equation}
  \frac{dR_\mu}{dt} \approx 0, \\ \forall \mu.
\end{equation}
Using Eq.\eqref{eq: differential eq for resources}, we get an explicit value for the resources:
\begin{equation}
  R_\mu \approx \frac{l_\mu+\sum_j \alpha_{\mu j}S_j}{m_\mu + \sum_k \gamma_{k\mu}S_k}.
\end{equation}
This expression can be used in Eq.\eqref{eq: differential eq for species} to get an effective system which describes the dynamics of the $N_S$ consumers:
\begin{equation}
  \frac{dS_i}{dt} = \left(\sum_\nu \left(\frac{\sigma_{i\nu}\gamma_{i\nu}l_\nu}{m_\nu+\sum_k \gamma_{k\nu}S_k} - \alpha_{\nu i}\right) -d_i + \sum_{\nu j} \frac{\sigma_{i\nu}\gamma_{i\nu}\alpha_{\nu j}}{m_\nu+\sum_{k}\gamma_{k\nu}S_k}S_j \right) S_i.
\end{equation}
This can be rewritten in a more compact way:
\begin{equation}
  \frac{dS_i}{dt} = p_i(S) S_i + \sum_j M_{ij}(S)S_i S_j \label{eq: effective equations of evolution}
\end{equation}
with
\begin{equation}
    p_i(S) = -\left(d_i+\sum_{\nu}\alpha_{\nu i}\right) + \sum_\nu \frac{\sigma_{i\nu}\gamma_{i\nu}l_\nu}{m_\nu+\sum_k \gamma_{k\nu}S_k}\text{ and } M_{ij}(S)=\sum_{\nu}\frac{\sigma_{i\nu}\gamma_{i\nu}\alpha_{\nu j}}{m_\nu+\sum_{k}\gamma_{k\nu}S_k}.
\end{equation}
If we assume the species $S_k$ are not too far away from their equilibrium values\footnote{Note that this is very rarely true, especially in the context of the study of structural stability, where entire species sometimes die out.}, \ie
\begin{equation}
S_k \approx S^*_k \ \forall k,
\end{equation}
then using Eq.\eqref{eq: positive m_nu} we can simplify $p_i$. Indeed,
\begin{equation}
m_\nu + \sum_k \gamma_{k\nu} S_k \approx m_\nu + \sum_k \gamma_{k\nu}S^*_k = \frac{l_\nu + \sum_k \alpha_{\nu k}S^*_k}{R^*_\nu} \label{eq: equality fluxes resource}
\end{equation}
Hence, the explicit dynamical dependence on $S$ can be removed from $p_i$ and $M_{ij}$:
\begin{equation}
p_i(S) \approx p_i \equiv - \left(d_i + \sum_\nu \alpha_{\nu i}\right) + \sum_\nu \frac{\sigma_{i\nu}\gamma_{i\nu}l_\nu R^*_\nu}{l_\nu + \sum_k \alpha_{\nu k}S^*_k},
\end{equation} and
\begin{equation}
M_{ij}(S) \approx M_{ij} \equiv \sum_\nu \frac{\sigma_{i\nu} \gamma_{i\nu} R^*_\nu \alpha_{\nu j}}{l_\nu + \sum_k{\alpha_{\nu k} S^*_k}}.
\end{equation}
\subsubsection{Perturbation analysis}
We study a system that we put close to an equilibrium $S^*$, \ie
\begin{equation}
S=S^*+\Delta S, \\ \text{with } \Delta S \ll 1.
\end{equation}
Written this way, the effective equations of motion Eq.\eqref{eq: effective equations of evolution} are equivalent to:
\begin{equation}
\frac{d\Delta S_i}{dt} = p_i(S^*+\Delta S)\left(S^*_i + \Delta S_i\right)+\sum_j M_{ij}(S^*+\Delta S)\left(S^*_i +\Delta S_i\right)\left(S^*_j +\Delta S_j\right).
\end{equation}
Since the deviations from equilibrium $\Delta S_i \ll 1$, we can forget the terms in higher power than quadratic:
\begin{equation}
\frac{d\Delta S_i}{dt} = \tilde{p}_i \Delta S_i + \sum_j E_{ij} \Delta S_j + \bigO(\Delta S^2),\label{eq: effective equ evol at O(D2)}
\end{equation}
with
\begin{equation}
\tilde{p}_i \equiv p_i(S^*) + \sum_k M_{ik}(S^*)S_k^*, \label{eq: tilde p effective system}
\end{equation}
and
\begin{equation}
E_{ij} \equiv \left(\partiald{p_i}{S_j}\evaluatedat{S^*}+M_{ij}(S^*)+\sum_k \partiald{M_{ik}}{S_j}\evaluatedat{S^*}S^*_k\right)S^*_i.
\end{equation}
After some computations, we can get $\tilde{p}_i$ and $E_{ij}$ in terms of the initial parameters. Indeed,
\begin{equation}
p_i(S^*)= -\left(d_i + \sum_\nu \alpha_{\nu i}\right) + \sum_\nu \frac{\sigma_{i\nu}\gamma_{i\nu}l_\nu}{m_\nu + \sum_k \gamma_{k\nu}S^*_k}
\end{equation}
and
\begin{equation}
M_{ik}(S^*) = \sum_\nu \frac{\sigma_{i\nu}\gamma_{i\nu}\alpha_{\nu j}}{m_\nu + \sum_k \gamma_{k\nu}S^*_k}.
\end{equation}
Hence, using Eq.\eqref{eq: tilde p effective system}:
\begin{equation}
\tilde{p}_i = - \left(d_i + \sum_\nu \alpha_{\nu i}\right) + \sum_\nu \frac{\sigma_{i\nu}\gamma_{i\nu}}{m_\nu + \sum_k \gamma_{k\nu}S^*_k}\left(l_\nu+\sum_{j}\alpha_{\nu j} S^*_j\right).
\end{equation}
This can be simplified using Eq.\eqref{eq: equality fluxes resource} and Eq.\eqref{eq: equilibrium species}:
\begin{equation}
\tilde{p}_i=-d_i +\sum_\nu \sigma_{i\nu}\gamma_{i\nu}R^*_\nu = \sum_\nu \alpha_{\nu i}.
\end{equation}
With a similar computation, one finds
\begin{equation}
E_{ij}=\sum_\nu \frac{\sigma_{i\nu}\gamma_{i\nu}S^*_i}{m_\nu+\sum_k \gamma_{k\nu}S^*_k} \left(\alpha_{\nu j}-\gamma_{j\nu}R^*_\nu\right).
\end{equation}
Finally, Eq.\eqref{eq: effective equ evol at O(D2)} can be recast in
\begin{equation}
\frac{d\Delta S_i}{dt} = \sum_j (J_E)_{ij} \Delta S_j,
\end{equation}
where the effective $N_S\times N_S$ jacobian matrix $J_E$ is defined by:
\begin{equation}
(J_E)_{ij}=\sum_\nu \left[\frac{\sigma_{i\nu}\gamma_{i\nu}S^*_i}{m_\nu+\sum_k \gamma_{k\nu}S^*_k} \left(\alpha_{\nu j}-\gamma_{j\nu}R^*_\nu\right)+\alpha_{\nu i}\delta_{ij}\right].
\end{equation}
We see that we without surprise we find again the $\Beta, \Gamma $ and $\Delta$ matrices coming from the jacobian at equilibrium:
\begin{equation}
\left(J_E\right)_{ij}=\sum_\nu \left[\frac{\Beta_{i\nu}\Gamma_{\nu j}}{\Delta_\nu}+\alpha_{\nu i} \delta_{ij}\right]
\end{equation}

This matrix determines the stability of the equilibrium. Namely if the largest eigenvalue of $J_E$ is positive, the equilibrium is unstable. If it is negative, the equilibrium is stable. If it is zero, the equilibrium is marginal.

\end{document}
